Statistical Inference of Transport Mechanisms and Long Time Scale Behavior from Time Series of Solute Trajectories in Nanostructured Membranes.

### 2020-03-10 Ben_Manuscripts/hdphmm/Outline.tex
Adding citations and beef to intro

  Lyotropic liquid crystals (LLC) are a class of amphiphilic molecules whose ordered phases
  can be cross-linked into mechanically strong membranes capable of highly selective
  separations.
  \begin{itemize}
    \item The shape of the LLC monomers and water content dictates the ordered phase 
    that they form. There are two phases of particular interest for membrane applications.
  	\item H\textsubscript{II} phase lyotropic liquid crystals are characterized by 
  	hexagonally packed, straight, pores while the Q\textsubscript{I} phase consists of
  	a tortuous network of 3D interconnected pores. 
  	\item In both cases the pores are uniform in size with radii on the order of 1 nm
         %MRS2: is it really a molecular weight cutoff?
  	giving them a very strict molecular weight cut-off.
	\item Additionally, they have the potential to disrupt conventional membrane separation
	techniques by being selective based not only on size and charge, but on chemical
	functionality as well.
	\item Their pores are lined with LLC monomer functional groups which can potentially
	be designed to interact with solutes in a chemically-specific manner
  \end{itemize}

@@

  Hidden Markov models (HMMs) are a useful and widely used technique
  for modeling sequences of observations where the probability of the next observation
  in a sequence depends only on a previous unobserved, latent or hidden, state.~\cite{beal_infinite_2002}
  %MRS2: it can depend on other things, but ALSO depends on the latent state.
  \begin{itemize}
    \item In the context of our simulations, the observations correspond to 
    the center of mass coordinates of the solutes versus time, and the states
    correspond to the dynamical behavior which give rise to those types
    of observations.
    \item Unfortunately, standard HMMs require the number of hidden states to be known
    a priori.
    \item One can partially overcome this by testing a range of numbers of 
    hidden states and determining which is the best representation of the
    data.
  \end{itemize}

@@

  We applied the IHMM algorithm to the radial and axial coordinates of each 
  solute center-of-mass trajectory.
  \begin{itemize}
    \item We measured the radial coordinate as the distance from the closest
    pore center.
    \item Because the pores are somewhat tortuous, we approximated the pore 
    center as a spline. See previous work.
    %MRS2: approximated the pore center line with a spline?
  \end{itemize}

@@

  %MRS2: add GAANN and PRF here to cover bases.
  Molecular simulations were performed using the Extreme Science and
  Engineering Discovery Environment (XSEDE), which is supported by National
  Science Foundation grant number ACI-1548562. Specifically, it used the Bridges
  system, which is supported by NSF award number ACI-1445606, at the Pittsburgh
  Supercomputing Center (PSC). This work also utilized the RMACC Summit supercomputer,
  which is supported by the National Science Foundation (awards ACI-1532235 and
  ACI-1532236), the University of Colorado Boulder, and Colorado State
  University. The Summit supercomputer is a joint effort of the University of
  Colorado Boulder and Colorado State University.

>>>>

  Lyotropic liquid crystals (LLC) are a class of amphiphilic molecules whose ordered phases
  can be cross-linked into mechanically strong membranes capable of highly selective
  separations.
  \begin{itemize}
    \item The shape of the LLC monomers and water content dictates the ordered phase 
    that they form. There are two phases of particular interest for membrane applications.
  	\item H\textsubscript{II} phase lyotropic liquid crystals are characterized by 
  	hexagonally packed, straight, pores while the Q\textsubscript{I} phase consists of
  	a tortuous network of 3D interconnected pores. 
  	\item In both cases the pores are uniform in size with radii on the order of 1 nm
    %MRS2: is it really a molecular weight cutoff?
    %BJC2: is molecular size better? Obviously there are other factors like shape, but I don't want to get into that
  	giving them a very strict molecular size cut-off.
	\item Additionally, they have the potential to disrupt conventional membrane separation
	techniques by being selective based not only on size and charge, but on chemical
	functionality as well.
	\item Their pores are lined with LLC monomer functional groups which can potentially
	be designed to interact with solutes in a chemically-specific manner
  \end{itemize}

@@

  Hidden Markov models (HMMs) are a useful and widely used technique
  for modeling sequences of observations where the probability of the next observation
  in a sequence depends on a previous unobserved, latent or hidden, state.~\cite{beal_infinite_2002}
  %MRS2: it can depend on other things, but ALSO depends on the latent state.
  \begin{itemize}
    \item In the context of our simulations, the observations correspond to 
    the center of mass coordinates of the solutes versus time, and the states
    correspond to the dynamical behavior which give rise to those types
    of observations.
    \item Unfortunately, standard HMMs require the number of hidden states to be known
    a priori.
    \item One can partially overcome this by testing a range of numbers of 
    hidden states and determining which is the best representation of the
    data.
  \end{itemize}

@@

  We applied the IHMM algorithm to the radial and axial coordinates of each 
  solute center-of-mass trajectory.
  \begin{itemize}
    \item We measured the radial coordinate as the distance from the closest
    pore center.
    \item Because the pores are somewhat tortuous, we approximated the pore 
    center with a spline. See previous work.
    %MRS2: approximated the pore center line with a spline?
  \end{itemize}

@@

  %MRS2: add GAANN and PRF here to cover bases.
  This work was supported in part by the ACS Petroleum Research Fund
  grant \#59814-ND7 and the Graduate Assistance in Areas of National Need (GAANN) 
  fellowship which is funded by the U.S. Department of Education. 
  Molecular simulations were performed using the Extreme Science and
  Engineering Discovery Environment (XSEDE), which is supported by National
  Science Foundation grant number ACI-1548562. Specifically, it used the Bridges
  system, which is supported by NSF award number ACI-1445606, at the Pittsburgh
  Supercomputing Center (PSC). This work also utilized the RMACC Summit supercomputer,
  which is supported by the National Science Foundation (awards ACI-1532235 and
  ACI-1532236), the University of Colorado Boulder, and Colorado State
  University. The Summit supercomputer is a joint effort of the University of
  Colorado Boulder and Colorado State University.
